---
title: 'Homework #10'
author: "Wenqiang Feng"
date: "3/6/2017"
output: html_document
---

### required library
```{r}
library(plyr)
```


### problem 24.12 a

- Load data
```{r}
dt <- read.table("./data/CH24PR12.txt", 
                col.names =  c("time", "gender", "sequence", "experience", "observation"))
dt[, 2] = as.factor(dt[,2])
dt[, 3] = as.factor(dt[,3])
dt[, 4] = as.factor(dt[,4])
```

- residual for ANOVA model
```{r}
fit = aov(time~gender+sequence+experience+gender*sequence+gender*experience+sequence*experience+gender*sequence*experience, data=dt)
residuals = cbind(dt, fit$residuals)
residuals
```



### Problem 24.12 b
- Calculate correlation coefficient
```{r }
res = lm(time~gender+sequence+experience+gender*sequence+gender*experience+sequence*experience+gender*sequence*experience, data=dt)
StdErr = summary(res)$sigma
n = length(res$residuals)
ExpVals = sapply(1:n, function(k) StdErr*qnorm((k-.375)/(n+.25)) )
r = cor(ExpVals, sort(res$residuals))
r
```


### problem 24.13 abcde

```{r eval=F, echo=F}
## Problem 24.12 a
### Calculate residuals
fit = aov(time~gender+sequence+experience+gender*sequence+gender*experience+
            sequence*experience+gender*sequence*experience, data=dt)
residuals = cbind(dt[,2:4], fit$residuals)
residuals
```






## Problem 24.13 a
### Calculate _estimated treatment means_

```{r}
estimated_treatment_mean = ddply(dt, .(gender, sequence, experience), 
                                 summarize, mean = mean(time))
estimated_treatment_mean
```

### AB plots
* No interaction effects 
* Both gender and sequence have a main effect.

```{r}
dat = dt[dt$experience == "1", ]
sequence = dat$sequence
gender = dat$gender
time = dat$time
interaction.plot(gender, sequence, time, main="Experience < 18 months")

## experience level = 1
dat = dt[dt$experience == "2", ]
sequence = dat$sequence
gender = dat$gender
time = dat$time
interaction.plot(gender, sequence, time, main="Experience > 18 months")

```


## Problem 24.13 b

```{r}
fit = aov(time~gender+sequence+experience+gender*sequence+gender*experience+
            sequence*experience+gender*sequence*experience, data=dt)
variance_table = summary(fit)
variance_table = round(as.matrix(variance_table[[1]]), 3)
variance_table[,1:3]
```

## Problem 24.13 c

### Test for three-factor interactions

* alternatives
    + H0: all $(\alpha\beta\gamma){ijk}$  equal zero
    + Ha: not all $(\alpha\beta\gamma){ijk}$ equal zero

* decision rules: if $F^*$ ≤ F(1-0.05; 2, 48), conclude H0; otherwise, conclude Ha. 
* conclusion: $F^*$ =0.011 < F(1-0.05; 2, 48) = 0.531, conclude H0.
* pvalue = 0.989

```{r}
variance_table[7,, drop=F]
```

## Problem 24.13 d

### Test for _AB_, _AC_, _BC_ interactions
#### AB interaction
* alternatives
    + H0: all $(\alpha\beta)_{ij}$  equal zero
    + Ha: not all $(\alpha\beta)_{ij}$ equal zero
* decision rules: if $F^*$ ≤ F(1-0.05; 2, 48), conclude H0; otherwise, conclude Ha. 
* conclusion: $F^*$ = 0.316 < F(1-0.05; 2, 48) = 3.190727, conclude H0.
* pvalue = 0.730
```{r}
qf(0.95,2,48)
variance_table[4,, drop=F]
```

#### AC interaction
* alternatives
    + H0: all $(\alpha\gamma)_{ik}$  equal zero
    + Ha: not all (αγ)ik  equal zero
* decision rules: if $F^*$ ≤ F(1-0.05; 1, 48), conclude H0; otherwise, conclude Ha. 
* conclusion: $F^*$ = 0.106 < F(1-0.05; 1, 48) = 4.042652, conclude H0.
* pvalue = 0.746
```{r}
qf(0.95,1,48)
variance_table[5,, drop=F]
```

#### BC interaction
* alternatives
    + H0: all $(\beta\gamma)_{jk}$  equal zero
    + Ha: not all $(\beta\gamma)_{jk}$  equal zero
* decision rules: if $F^*$ ≤ F(1-0.05; 2, 48), conclude H0; otherwise, conclude Ha. 
* conclusion: $F^*$ = 0.106 < F(1-0.05; 2, 48) = 0.531, conclude H0.
* pvalue = 0.591
```{r}
qf(0.95,2,48)
variance_table[6,, drop=F]
```


## Problem 24.13 e
### Test for main effects

#### main effect from A
* alternatives
    + H0: all $\alpha_i$ (i=1,2)  equal zero
    + Ha: not all $$\alpha_i$ $ equal zero
* decision rules: if $F^*$ ≤ F(1-0.05; 1, 48), conclude H0; otherwise, conclude Ha. 
* conclusion: $F^*$ = 629.76 > F(1-0.05; 1, 48) =4.042652, conclude Ha.
* pvalue = 0
```{r}
qf(0.95,1,48)
variance_table[1,, drop=F]
```

#### main effect from B
* alternatives
    + H0: all $\beta_j$ (j=1,2,3)  equal zero
    + Ha: not all $\beta_j$ equal zero
* decision rules: if $F^*$ ≤ F(1-0.05; 2, 48), conclude H0; otherwise, conclude Ha. 
* conclusion: $F^*$ = 28.74 > F(1-0.05; 2, 48) = 3.190727, conclude Ha.
* pvalue = 0
```{r}
qf(0.95,2,48)
variance_table[2,, drop=F]
```

#### main effect from C
* alternatives
    + H0: all $\gamma_k$ (k=1,2)  equal zero
    + Ha: not all $\gamma_k$ equal zero
* decision rules: if $F^*$ ≤ F(1-0.05; 1, 48), conclude H0; otherwise, conclude Ha. 
* conclusion: $F^*$ = 445.668 > F(1-0.05; 1, 48) = 4.042652, conclude Ha.
* pvalue = 0
```{r}
qf(0.95,1,48)
variance_table[3,, drop=F]
```



## Problem 24.14 a

D1 = 189.8
```{r}
D = ddply(dt, .(gender), summarize, mean = mean(time))
D1 = D[1,2] - D[2,2]
D1
```

D2 = -57.25
```{r}
D = ddply(dt, .(sequence), summarize, mean = mean(time))
D2 = D[1,2] - D[2,2]
D2
```

D3 = 6.6
```{r}
D = ddply(dt, .(sequence), summarize, mean = mean(time))
D3 = D[1,2] - D[3,2]
D3
```


D4 = 63.85
```{r}
D = ddply(dt, .(sequence), summarize, mean = mean(time))
D4 = D[2,2] - D[3,2]
D4
```

D5 = 159.6667
```{r}
D = ddply(dt, .(experience), summarize, mean = mean(time))
D5 = D[1,2] - D[2,2]
D5
```

MSE = 858.042
```{r}
MSE = variance_table[8,3,drop=F]
MSE
```


* a = 2
* b = 3
* c = 2
* n = 5

* s{D1} = ((MSE/(nbc))[1^2 + (-1)^2])^0.5 = 7.563253
* s{D2} = s{D3} = s{D4} = ((MSE/(nac))[1^2 + (-1)^2])^0.5 = 9.263056
* s{D5} = ((MSE/(nab))[1^2 + (-1)^2])^0.5 = 7.563253
* B = t[1-a/2g; (n-1)\*abc] = t(1-0.1/(2\*5); 48) = t(0.99; 48) = 2.406581

* confidence coefficients
    + D1 = [171.5984, 208.0016]
    + D2 = [-79.54229, -34.95771]
    + D3 = [-15.69229, 28.89229]
    + D4 = [41.55771, 86.14229]
    + D5 = [141.4651, 177.8683]
* findings: D1, D2, D4, D5 are significant, but D3 is not significant.


## Problem 24.14 b

* Y231 = 1020.4
```{r}
Y = ddply(dt, .(gender, sequence, experience), summarize, mean = mean(time))
Y213 = Y[11,]
Y213
```

* s{Y231} = (MSE/n)^0.5 = 13.09994
* t(1-0.05/2; 48) = 2.010635
* confidence interval = [Y - t(1-0.05/2; 48)\*s{Y231}, Y + t(1-0.05/2; 48)\*s{Y231}]
* confidence interval = [994.0608, 1046.739]

