---
title: 'Homework #11'
author: "Wenqiang Feng"
date: "3/25/2017"
output: html_document
---


## require package 
```{r message=FALSE, warning=FALSE}
library(plyr)
library(asbio)
```

## Load data
```{r}
dt <- read.table("./data/CH22PR11.txt", 
                col.names =  c("physical_therapy_days", "fitness_status", "obs", "age"))

dt$fitness_status = as.factor(dt$fitness_status)
dt$obs = as.factor(dt$obs)
```

## Problem a
### Test residuals normality

* W = 0.98277, p-value = 0.9409
* Normality test indicates that the residuals are normally distributed

```{r}
fit = aov(physical_therapy_days~age+fitness_status, dt)
shapiro.test(fit$residuals)
```

## Problem b
### interaction between treatment factor and covariate

* F_star = 0.3359, P value = 0.719 
* So no interaction between factor and covariate

```{r}
fit.interaction = aov(physical_therapy_days~age+fitness_status+age:fitness_status, dt)
summary(fit.interaction)

variance_table = summary(fit.interaction)[[1]]
variance_table[3,,drop=F]
```

* Plot below also indicates no interaction effect

```{r}
plot(dt$physical_therapy_days~dt$age, col=dt$fitness_status, pch=18)
```

## Problem c
### test for relationship between covariate and response variable
* The covariate is significantly correlated with the response variable. 

```{r}
cortest = cor.test(dt$physical_therapy_days,dt$age)
cortest
```

* correlation coefficient = 0.8764434, p-value = 1.972299e-08
```{r}
cortest$estimate
cortest$p.value
```

## Problem d
### test for treatment effect

* There is a significant treatment effect on the physical therapy days
* F_star = 372.61, p-value = 2.257e-15
```{r}
fit.treatment = aov(physical_therapy_days~age+fitness_status, dt)
variance_table = summary(fit.treatment)[[1]]
variance_table
variance_table[2,,drop=F]
```


## Problem e
### pairwise comparisons between treatment levels

* All pairwise comparisons show significant difference.

```{r}
fit.tukey = aov(physical_therapy_days~fitness_status, dt)
TukeyHSD(fit.tukey, conf.level = 0.95)
boxplot(physical_therapy_days ~ fitness_status, dt, xlab="Treatment levels", 
        ylab = "physical therapy days")
```