---
title: "Homework#12"
author: "Wenqiang Feng"
date: "4/2/2017"
output: html_document
---


## require package 
```{r message=FALSE, warning=FALSE}
library(plyr)
library(asbio)
```

## Load data
```{r}
dt <- read.table("./data/CH27PR06.txt", 
                col.names =  c("sales", "stores", "price"))
dt$stores = as.factor(dt$stores)
dt$price = as.factor(dt$price)
```


## Problem 27.6 a
### Residuals and normality test

* The residual is normally distributed

```{r}
fit = aov(sales~stores+price, dt)
residuals = matrix(fit$residuals, nrow=8, byrow=T)
shapiro.test(residuals)
```

## Problem 27.6 b
### Aligned residual dot plots

* Yes. The plot supports the assumption of constancy of the error variance

```{r}
plot(as.numeric(dt$price)~fit$residuals, xlab="Residuals", ylab="Price", col=dt$price, pch=16)
```


## Problem 27.6 c
### Interaction between subjects and treatments

* Yes. the assumption of no interaction between subjects and treatments is reasonable.

```{r}
stores = dt$stores
sales = dt$sales
price = dt$price
interaction.plot(stores, price, sales)
```

## Problem 27.7 a
### Analysis of variance table

```{r}
fit = aov(sales~stores+price, dt)
variance_table = round(summary(fit)[[1]], 3)
variance_table
```

## Problem 27.7 b
### Whether or not the mean sales of grapefruits differ

* alternatives
    + H0: all Tj equal zero
    + Ha: not all Tj equal zero

* decision rules: if F\_star <= F(.95; 2, 14), conclude H0; otherwise, conclude Ha. 
* conclusion: F\_star = 49.346 > F(.95; 2, 14) = 3.738892, conclude Ha.
* pvalue < 2.2e-16

```{r}
variance_table
variance_table[2,,drop=F]
```

## Problem 27.7 c
### Pairwise comparisons of the price level means

* Tukey procedure was applied for the pairwise comparisons. 
* All differences are significant
  + The mean of level 1 is the largest and the mean of level 3 is the smallest.
  + the differnce between level 1 and level 3 is the largest.
* Below is the results:

```{r}
TukeyHSD(fit, "price", ordered=T)
```