---
title: "Homework#13"
author: "Wenqiang Feng"
date: "4/2/2017"
output: html_document
---

## require package 
```{r message=FALSE, warning=FALSE}
library(plyr)
```

## Load data
```{r}
dt <- read.table("./data/CH27PR13.txt", 
                col.names =  c("response", "store", "display", "time_period"))
dt$store = as.factor(dt$store)
dt$display = as.factor(dt$display)
dt$time_period = as.factor(dt$time_period)
```

## Problem 27.13 a
* obtain residuals 

```{r}

fit = aov(response~display * time_period + Error(store/(display)), dt)
residuals = proj(fit)[[4]][, "Residuals"]
residuals = data.frame(id=dt$display, res=residuals)
residuals = arrange(residuals, id)
residuals = matrix(residuals$res, 8, 4, byrow=T)
colnames(residuals) =  c("k=1", "k=2", "k=3", "k=4")
rownames(residuals) = paste("i", c(1:4, 1:4), sep="=")
residuals
```

* residuals against the fitted values

```{r}
plot(proj(fit)[[4]][, "Residuals"]~dt$response-proj(fit)[[4]][, "Residuals"], 
     xlab="fitted values", ylab="residuals", ylim=c(-30, 30))
abline(h=0)
```

* qqplot

```{r echo=F }
qqplotData = qqnorm(proj(fit)[[4]][, "Residuals"], plot.it = FALSE)
```

```{r}
qqnorm(qqplotData$y, ylim=c(-30, 30))
qqline(qqplotData$y)
```

* Based on plots above, the model is appropriate to explain this data.


## Problem 27.13 b

* Based on the plot, the model is appropriate here.

```{r}
par(mfrow=c(1,2))
dat = subset(dt, display==1)
store = dat$store
time_period = dat$time_period
response = dat$response
interaction.plot(time_period, store, response, main="display 1")

dat = subset(dt, display==2)
store = dat$store
time_period = dat$time_period
response = dat$response
interaction.plot(time_period, store, response, main="display 2")
```


## Problem 27.15 a

```{r}
S_display = summary(fit)[[1]][[1]][1,] + summary(fit)[[2]][[1]][2,]
S_display[1,3] = S_display[1,2]/S_display[1,1]
S_display[1,4] = S_display[1,3]/summary(fit)[[3]][[1]][3,3]
S_display[1,5] = 0
variance_table = rbind(summary(fit)[[2]][[1]][1,], S_display,  summary(fit)[[3]][[1]])
rownames(variance_table) = c("display", "S(display)", "time period", 
                             "display:time period", "Error")
variance_table
```

## Problem 27.15 b

* No interaction effects
* Main effects exist

```{r}
meanTime = ddply(dt, .(time_period, display), summarise, mean=mean(response))
dt <- read.table("./data/CH27PR13.txt", 
                col.names =  c("response", "store", "display", "time_period"))
plot(subset(dt, time_period==1&display==1)$time_period, subset(dt, time_period==1&display==1)$response, 
     xlim=c(0,5), ylim=c(0, 1200), xlab="time period", ylab = "sales", pch=16, col=2)
points(subset(dt, time_period==2&display==1)$time_period, subset(dt, time_period==2&display==1)$response, 
       xlim=c(0,5), ylim=c(0, 1200), xlab="time period", ylab = "sales", pch=16, col=2)
points(subset(dt, time_period==3&display==1)$time_period, subset(dt, time_period==3&display==1)$response, 
       xlim=c(0,5), ylim=c(0, 1200), xlab="time period", ylab = "sales", pch=16, col=2)
points(subset(dt, time_period==4&display==1)$time_period, subset(dt, time_period==4&display==1)$response, 
       xlim=c(0,5), ylim=c(0, 1200), xlab="time period", ylab = "sales", pch=16, col=2)

mean1 = subset(meanTime, display==1)
points(mean1$mean~mean1$time_period, type='l', col=2, lwd=3)
points(subset(dt, time_period==1&display==2)$time_period, subset(dt, time_period==1&display==2)$response, 
       xlim=c(0,5), ylim=c(0, 1200), xlab="time period", ylab = "sales", pch=16, col=3)
points(subset(dt, time_period==2&display==2)$time_period, subset(dt, time_period==2&display==2)$response, 
       xlim=c(0,5), ylim=c(0, 1200), xlab="time period", ylab = "sales", pch=16, col=3)
points(subset(dt, time_period==3&display==2)$time_period, subset(dt, time_period==3&display==2)$response, 
       xlim=c(0,5), ylim=c(0, 1200), xlab="time period", ylab = "sales", pch=16, col=3)
points(subset(dt, time_period==4&display==2)$time_period, subset(dt, time_period==4&display==2)$response, 
       xlim=c(0,5), ylim=c(0, 1200), xlab="time period", ylab = "sales", pch=16, col=3)

mean2 = subset(meanTime, display==2)
points(mean2$mean~mean2$time_period, type='l', col=3, lwd=3)
text(4.5, 800, 'display 1')
text(4.5, 600, 'display 2')
```


## Problem 27.13 c

* alternatives
    + H0: all (alpha*beta)jk equal zero
    + Ha: not all (alpha*beta)jk equal zero

* decision rules: if F\_star <= F(.975; 3, 18), conclude H0; otherwise, conclude Ha. 
* conclusion: F\_star = 1.083 < F(.975; 3, 18) = 3.953863, conclude H0.
* pvalue = 0.381

```{r}
summary(fit)[[3]][1]
```

## Problem 27.13 d

* alternatives
    + H0: all (alpha)j equal zero
    + Ha: not all (alpha)j equal zero

* decision rules: if F\_star <= F(.975; 1, 6), conclude H0; otherwise, conclude Ha. 
* conclusion: F\_star = 1.083 < F(.95; 1, 6) = 8.813101, conclude H0.
* pvalue = 0.3813 

```{r}
summary(fit)[[3]][[1]][2,]
```


* alternatives
    + H0: all (beta)j equal zero
    + Ha: not all (beta)j equal zero

* decision rules: if F\_star <= F(.975; 1, 6), conclude H0; otherwise, conclude Ha. 
* conclusion: F\_star = 83.636 > F(.95; 1, 6) = 8.813101, conclude Ha.
* pvalue = 9.22e-11

```{r}
summary(fit)[[3]][[1]][1,4, drop=FALSE]
```



## Problem 27.15 e
* L1 = 182.375
* L2 = -9.375
* L3 = 20.625
* L4 = -103.375
* s = 4
* a = 2
* b = 4
* s{Li} = 2MSB.S(A)/as = ((2\*213)/(2\*4))^.5 = 7.29726
* s{L1} = 2MSS(A)/bs = ((2\*213)/(2\*4))^.5 = 250.1674
* Bi = t(.9875; 18) = 2.445006
* B1 = t(.9875, 6) = 2.968687
* Confidence intervals
    + L1 = [L1-s{L1}B1, L1+s{L1}B1] = [-560.2937, 925.0437], no significant difference
    + L2 = [L2-s{Li}Bi, L2+s{Li}Bi] = [-27.21684, 8.466844], no significant difference
    + L3 = [L3-s{Li}Bi, L3+s{Li}Bi] = [2.783156, 38.46684], significant difference
    + L4 = [L4-s{Li}Bi, L4+s{Li}Bi] = [-121.2168, -85.53316], significant difference
    
```{r}
L = ddply(dt, .(display), summarise, mu = mean(response))
L1 = L$mu[1] - L$mu[2]
L1

L = ddply(dt, .(time_period), summarise, mu = mean(response))
L2 = L$mu[1] - L$mu[2]
L2

L3 = L$mu[2] - L$mu[3]
L3

L4 = L$mu[3] - L$mu[4]
L4
```
