---
title: "Homework#15"
author: "Wenqiang Feng"
date: "4/16/2017"
output: html_document
---

## Required library
```{r message=FALSE, warning=FALSE}
library(sas7bdat)
library(effects)
library(pROC)
library(ROCR)
library(sjPlot)
```


## Load data set
```{r cache=TRUE}
dt <-read.sas7bdat("data/pva_raw_data.sas7bdat", debug=TRUE)
```

## Fit Logistic Regression Model

```{r}
fit.log <- glm(TARGET_B ~ PEP_STAR + RECENT_AVG_GIFT_AMT +       
                 FREQUENCY_STATUS_97NK, 
               family=binomial(link='logit'),data=dt)
```

```{r}
summary(fit.log)
```

```{r}
prob=predict(fit.log,type=c("response"))
dt$prob=prob
g <- roc(TARGET_B ~ prob, data = dt)
plot(g)   
```

## odds ratios and 95% CI
```{r}
exp(cbind(OR = coef(fit.log), confint(fit.log)))
```

## Effect plots with confidence bands 
```{r}
plot(allEffects(fit.log))
```

```{r}
sjp.glm(fit.log,
        xisLabels.y=lab,
        type="bars",
        gridBreaksAt=0.4)
```
## Interpretation

- Interpret the c statistic: The C-statistic is a measure of goodness of fit for binary outcomes in a logistic regression model. A value below 0.5 indicates a very poor model. A value of 0.5 means that the model is no better than predicting an outcome than random chance.
Values over 0.7 indicate a good model. Values over 0.8 indicate a strong model. A value of 1 means that the model perfectly predicts those group members who will experience a certain outcome and those who will not. From the Roc Curve, we may conclude that the model is no better than predicting on outcome than random chance.
- Interpret the odds ratio for PEP\_STAR: the coefficient of oadds ratio for PEP\_STAR is 1.3604636 which means, when PEP\_STAR increases 1 unit and the odds will increase 1.3604636 units when the other parameters are fixed.
- Interpret the effect plot for RECENT\_AVG\_GIFT\_AMT: the coefficient of oadds ratio for PEP\_STAR is 0.9944644 which means, when RECENT\_AVG\_GIFT\_AMT increases 1 unit and the odds will increase 0.9944644 units when the other parameters are fixed.
