---
title: "Homework #2"
author: "Ming Chen & Wenqiang Feng"
date: "2/2/2017"
output: html_document
---

## **1. Chapter 5 Problems 2**

#### Set matrix A, B, and C
```{r}
A <- matrix(c(2,1,3,5,5,7,4,8), ncol=2, byrow = TRUE)
A
B <- matrix(c(6,9,3,1), ncol=1, byrow = TRUE)
B
C <- matrix(c(3,8,8,6,5,1,2,4), ncol=2, byrow = TRUE)
C
```
#### (1) *$A + C$*
```{r}
A + C
```

#### (2) *$A - C$*
```{r}
A - C
```

#### (3) *$B'A$*
```{r}
t(B) %*% A
```

#### (4) *$AC'$*
```{r}
A %*% t(C)
```

#### (5) *$C'A$*
```{r}
t(C) %*% A
```

## **2. Chapter 5 Problems 5**   

### set data
```{r}
X = matrix(c(4,1,2,3,3,4), ncol=1,byrow = TRUE)
aX = cbind(1, X)
X
aX
Y = matrix(c(16,5,10,15,13,22), ncol=1, byrow = TRUE)
Y
```
#### (1) *$Y'Y$* 
```{r}
t(Y) %*% Y
```

#### (2) *$X'X$*
```{r}
t(aX) %*% aX
```

#### (3) *$X'Y$* 
```{r}
t(aX) %*% Y
```

## **3. Chapter 5 Problems 10**  
   
#### set data
```{r}
A = matrix(c(2,4,3,1), ncol=2, byrow = TRUE)
A
B = matrix(c(4,3,2,6,5,10,10,1,6), ncol=3, byrow = TRUE)
B
```

#### (1) *Inverse of matrices*

* *Inverse of matrix A: $A^{-1}$*
```{r}
A_inv = solve(A)
A_inv
```

* *Inverse of matrix B: $B^{-1}$ *
```{r}
B_inv = solve(B)
B_inv
```

#### (2) *Checking *

* checking for A
```{r}
A %*% A_inv
```
* checking for B
```{r}
B %*% B_inv
round(B %*% B_inv,digits=15)
```
## **4. Chapter 5 Problems 11**

#### set data
```{r}
A = matrix(c(5,4,1,1,0,9,3,5,6), ncol=3)
A
```

#### (1) Inverse of A
```{r}
A_inv = solve(A)
A_inv
```
#### (2) checking 

```{r}
A %*% A_inv
round(A %*% A_inv,digits=15)
```

### **4. Chapter 5 Problems 24**

### set data
```{r}
X = matrix(c(4,1,2,3,3,4), ncol=1,byrow = TRUE)
aX = cbind(1, X)
Y = matrix(c(16,5,10,15,13,22), ncol=1, byrow = TRUE)
```
#### a (1)
```{r}
betahat = solve(t(aX)%*%aX)%*%t(aX)%*%Y
betahat
```

#### a (2)
```{r}
Yhat = aX%*%betahat
residuals = Y - Yhat
residuals
```

#### a (3) $SSR = \sum (\hat{Y} - \bar{Y})^2$
```{r}
SSR = sum((Yhat - mean(Y))^2)
SSR
```

#### a (4) $SSE = \sum (Y - \hat{Y})^2$
```{r}
SSE = sum((Y-Yhat)^2)
SSE
```

#### a (5) *estimated variance-coveraiance matrix of b*
```{r}
sigma_squared = SSE/(length(X)- 1 - 1)
sigma_squared
sigma_squared * solve(t(aX)%*%aX)
```

#### a (6)
```{r}
betahat
```

```{r eval=FALSE}
Y = 0.439 + 4.610 * X
```

```{r}
X4 = 4
Yhat4 = 0.439 + 4.610 * X4
Yhat4
```

#### a (7) *$s^2$ {pred} when Xh = 4*
```{r}
MSE = SSE/(length(X)-2)
Xh = c(1, 4)
MSE * (1 + t(Xh)%*%solve(t(aX)%*%aX)%*%Xh)
```


### b *calculation*
```{r}
MSE = SSE/(length(X) - 2)
s_squared_b = MSE * solve(t(aX)%*%aX)
s_squared_b
```
#### b (1) *$s\{b_0,b_1\}$*
```{r}
s_squared_b[1,2]
```

#### b.(2) *$s^2\{b_0\}$*
```{r}
s_squared_b[1,1]
```

#### b.(3) *$s\{b_1\}$*
```{r}
(s_squared_b[2,2])^0.5
```

#### c. *matrix H*
```{r}
H = aX%*%solve(t(aX)%*%aX)%*%t(aX)
H
```


#### d. *$s^2\{e\}$*

```{r}
I = diag(1, length(X))
MSE = SSE/(length(X)- 1 - 1)
MSE * (I-H)
```