---
title: 'Homework #4'
author: "Wenqiang Feng"
date: "2/2/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **1. Chapter 7 Problems 7**

#### load data 
```{r}
library(car)
data618 <- read.table("./data/CH06PR18.txt", col.names = c("Y", "X1", "X2", "X3", "X4"))
```

#### Problem 7.7(a)

* $SSR(X_4)$: only variable X4 in the model
      - Method 1
```{r}
anova_x4 = Anova(lm(Y~ X4, data=data618))
SSR_X4 = anova_x4[1][1,1]
SSR_X4
```

      - Method 2
```{r}
SSR <- anova(lm(Y~X4, data=data618),lm(Y~1, data=data618))
SSR_X4 = SSR$RSS[2]-SSR$RSS[1]
SSR_X4
```

Hence, $SSR(X_4) = `r SSR_X4`$.

* $SSR(X_1|X_4)= SSR(X_1,X_4)-SSR(X_4).$ $X_1$ is added to the previous model
     
      - Method 1

```{r}
anova_x1_x4 = Anova(lm(Y~X4+X1, data=data618))
SSR_X1_X4 = anova_x1_x4[1][2,1,drop=FALSE]
SSR_X1_X4
```
      - Method 2

```{r}
RSS <- anova(lm(Y~X1+X4, data=data618),lm(Y~X4, data=data618))
SSR_X1_X4 = RSS$RSS[2]-RSS$RSS[1]
SSR_X1_X4
```
Hence, $SSR(X_1|X_4)= SSR(X_1,X_4)-SSR(X_4) = `r RSS$RSS[2] ` -`r RSS$RSS[1]`= `r SSR_X1_X4`$.

* $SSR(X_2|X_1, X_4) =SSR(X_1, X_2, X_4)-SSR(X_1,X_4)$

      - Method 1
```{r}
anova_x2_x1x4 = Anova(lm(Y~X4+X1+X2, data=data618))
SSR_X2_X1X4 = anova_x2_x1x4[1][3,1,drop=FALSE]
SSR_X2_X1X4
```
      - Method 2

```{r}
RSS = anova(lm(Y~X4+X1+X2, data=data618),lm(Y~X1+X4, data=data618))
SSR_X2_X1X4 = RSS$RSS[2]-RSS$RSS[1]
SSR_X2_X1X4
```

Hence, $SSR(X_2|X_1, X_4) =SSR(X_1, X_2, X_4)-SSR(X_1,X_4) = `r RSS$RSS[2]`- `r RSS$RSS[1]`= `r SSR_X2_X1X4`$.

* $SSR(X_3|X_1, X_2, X_4) = SSR(X_1,X_2,X_3,X_4)-SSR(X_1,X_2,X_4)$

      - Method 1
```{r}
anova_x3_x1x4x2 = Anova(lm(Y~X4+X1+X2+X3, data=data618))
SSR_X3_X1X2X4 = anova_x3_x1x4x2[1][4,1,drop=FALSE]
SSR_X3_X1X2X4
```
        - Method 2
        
```{r}
RSS = anova(lm(Y~X1+X2+X3+X4, data=data618),lm(Y~X1+X2+X4, data=data618))
SSR_X3_X1X2X4 = RSS$RSS[2]-RSS$RSS[1]
SSR_X3_X1X2X4
```

Hence, $SSR(X_3|X_1, X_2, X_4) = SSR(X_1,X_2,X_3,X_4)-SSR(X_1,X_2,X_4) = `r RSS$RSS[2]`- `r RSS$RSS[1]`= `r SSR_X3_X1X2X4`$


#### Problem 7.7(b)
* Hypothesis
    - $H_0$: $\beta_3$ = 0, $H_{\alpha}$: not all $\beta_3\neq 0$
    - test statistic ($\alpha =0.01, n=81, p=5$):
\[ F^* = \frac{SSR(X_q,\cdots,X_{p-1}|X_1,\cdots, X_{q-1})}{p-q}\div \frac{SSE(X_1,\cdots,X_{p-1})}{n-p}\]
         - If $F^* \leq F(1-\alpha; p-q, n-p) = F(0.99, 1, 76)$, conclude $H_0$, 
         - If $F^* > F(1-\alpha; p-q, n-p) = F(0.99, 1, 76)$  conclude $H_a$.

* $F^*$ and $F$  
Since 
      \[ F^* = \frac{`r SSR_X3_X1X2X4`}{5-4}\div \frac{`r RSS$RSS[1]`}{81-5}= `r SSR_X3_X1X2X4/(RSS$RSS[1]/(81-5))`.\]
And     
\[ F = F(1-\alpha, p-q, n-p) = F(0.99,1,76) = `r qf(0.99, 1,76)` \]

$F^* < F$, hence we conclude $H_0$.  

* p-value 
```{r}
pvalue = pf(q=0.3247534, df1=1, df2=76, lower.tail=FALSE)
pvalue
```

Hence the p-value from the F test socre  `r SSR_X3_X1X2X4/(RSS$RSS[1]/(81-5))` is `r pvalue` .




## **2. Chapter 7 Problems 8**

* Hypothesis
    - $H_0$: $\beta_2=\beta_3$ = 0, $H_{\alpha}$: not both $\beta_i$ = 0 (i=2,3)
    - test statistic ($\alpha =0.01, n=81, p=5$):
 \[ F^* = \frac{SSR(X_q,\cdots,X_{p-1}|X_1,\cdots, X_{q-1})}{p-q}\div \frac{SSE(X_1,\cdots,X_{p-1})}{n-p}\]
         - If $F^* \leq F(1-\alpha; p-q, n-p) = F(0.99, 2, 76)$, conclude $H_0$, 
         - If $F^* > F(1-\alpha; p-q, n-p) = F(0.99, 2, 76)$  conclude $H_a$.

* $F^*$ and $F$ 

$SSR(X_3|X_1, X_2, X_4)$ = 
```{r}
SSR_X3_X1X2X4
```
$SSR(X_2|X_1, X_4)$ =  
```{r}
SSR_X2_X1X4
```
$SSR(X_2,X_3|X_1, X_4)= SSR(X2|X_1,X_4) +SSR(X_3|X_1,X_2,X_4)=$ 
```{r}
SSR_X2X3_X1X4 = SSR_X2_X1X4+SSR_X3_X1X2X4
SSR_X2X3_X1X4
```
$SSE(X_1,X_2,X_3,X_4)$ = 
```{r}
anova_x3_x1x4x2 = Anova(lm(Y~X4+X1+X2+X3, data=data618))
SSE = anova_x3_x1x4x2[1][5,1,drop=FALSE]
SSE
```

Since $SSR(X_2,X_3|X_1,X_4) = `r SSR_X2X3_X1X4`$ and $SSE(X_1,X_2,X_3,X_4)= `r SSE`$,


\[F^*= \frac{`r SSR_X2X3_X1X4`}{5-3}\div \frac{`r SSE`}{76}=`r (SSR_X2X3_X1X4/2)/(SSE/76)`\]
while 

```{r}
Fvalue = qf(0.99, 2,76)
Fvalue
```

Since $F^*> `r Fvalue`$, hence we conclude $H_\alpha$.

* Conclusion:
    + $F^*$ = 10.9389 > F(.99, 2, 76), conclude $H_\alpha$. Not both $\beta_2$ and $\beta_3$ = 0
    + Not both $X_2$ and $X_3$ can be dropped from the regression model given that $X_2$ and $X_4$ are retained
    
## **3. Chapter 7 Problems 10**

#### Hypothesis
* Alternatives:
    + $H_0$: $\beta_1$ = -0.1 and $\beta_2$ = 0.4
    + $H_\alpha$: not both equatities are true
    
* Decision rules ($df_R=81-3=78,df_F=81-5=76$):
    + if $F^* \leq F(1-\alpha,df_R-df_F, df_F)=F(.99, 2, 76)$, conclude $H_0$
    + if $F^* >  F(1-\alpha,df_R-df_F, df_F)= F(.99, 2, 76)$, conclude $H_\alpha$

#### Full model
\[ Y = \beta_0 + \beta_1\cdot X_1 + \beta_2\cdot X_2 +\beta_\cdot X_3+\beta_4\cdot X_4\]

#### Reduced model
* The reduced model $Y' = Y + 0.1\cdot X_1 - 0.4\cdot X_2$
* The data set for reduced model is
```{r}
Y = data618$Y + data618$X1*.1 - data618$X2*.4
data618_reduced = cbind(Y, data618[,c(4,5)])
data618_reduced
```

#### Calculate SSE(reduced model): SSE($X_1,X_4$) given known beta2 and beta3
```{r}
anova_reduced_model = Anova(lm(Y~X3+X4, data=data618_reduced))
anova_reduced_model
SSE_reduced_model = anova_reduced_model[1][3,1,drop=FALSE]
SSE_reduced_model
```

#### SSE(full model): $SSE(X_1,X_2,X_3,X_4)$ = 
```{r}
SSE
```

#### calculate $F^*$ 
```{r}
F_star = ((SSE_reduced_model-SSE)/2)/(SSE/76)
F_star[,,drop=TRUE]
```

#### calculate F(.99, 2, 76)
```{r}
F_2_76 = qf(.99, 2, 76)
F_2_76
```

#### Conclusion
* F_star = `r F_star[,,drop=TRUE]` < F(.99, 2, 76) = `r F_2_76`, conclude $H_0$: $\beta_1$ = -0.1 and $\beta_2$ = 0.4.
