---
title: "Homework #5"
author: "Wenqiang Feng"
date: "2/7/2017"
output: html_document
---


## **Homework on Variable Selection**

* required library in this homework
```{r warning = FALSE, message = FALSE}
library("R330")
library(MASS)
library(leaps)
```
#### load data 
```{r}
dataC07 <- read.table("./data/APPENC07.txt", col.names = c('ID',  'SalesPrice','FinishedSquareFeet','NofBedrooms','NofBathrooms',
'AirConditioning','GarageSize','Pool','YearBuilt',
'Quality','Style','LotSize','AdjacenttoHighway'))
```

### 1. Use All Possible Regressions to select a best model

*  run allpossregs()

```{r}
allPossRegsSelect = allpossregs(SalesPrice~.-Quality-Style, data=dataC07)
allPossRegsSelect
```

* Select the best model: The best model has the smallest Cp. Here it is model 9.
The variables that should be included in the model:
```{r}
minCp = which.min(allPossRegsSelect[,"Cp"])
minCp
vars = which(allPossRegsSelect[,-c(1:8)][minCp,] != 0)
vars 
```

* Obtain the model parameters
```{r}
allPossRegsFit = lm(SalesPrice ~ .-NofBathrooms-Pool , data = dataC07)
allPossRegsFit
```

### 2. Backward elimination

```{r}
fit0 = lm(SalesPrice~.-Quality-Style, data=dataC07)
BEfit = stepAIC(fit0, direction="backward")
coef(BEfit)
```

### 3. Forward selection

```{r}
fit0 = lm(SalesPrice~1, data=dataC07)
FSfit = stepAIC(fit0, scope=SalesPrice~FinishedSquareFeet+NofBedrooms+NofBathrooms+
                AirConditioning+GarageSize+Pool+YearBuilt+
                LotSize+AdjacenttoHighway, direction="forward")
coef(FSfit)
```

### 4. Stepwise

```{r}
fit0 = lm(SalesPrice~.-Quality-Style, data=dataC07)
SWfit = stepAIC(fit0, direction="both")
coef(SWfit)
```


### 5. Maximum R squared improvement

```{r}
fit0 = lm(SalesPrice~1, data=dataC07)
maxRfit = add1(fit0, scope=SalesPrice~FinishedSquareFeet+NofBedrooms+NofBathrooms+AirConditioning+GarageSize+Pool+YearBuilt+LotSize+AdjacenttoHighway, test="F")
maxRfit
```

### 6. Using the information from the different variable selection methods, fit the best model.

```{r}
bestFit = lm(SalesPrice~FinishedSquareFeet+ NofBedrooms+
               AirConditioning+GarageSize+YearBuilt+LotSize,
               data=dataC07)
coef(bestFit)
```
