---
title: "Homework #6"
author: "Wenqiang Feng"
date: "2/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## *Homework on Identification of Influential Observations*

* required library in this homework
```{r warning = FALSE, message = FALSE}
library(dplyr)
```
#### load data 
```{r}
dataC07 <- read.table("./data/APPENC07.txt", col.names = c('ID',  
                  'SalesPrice','FinishedSquareFeet','NofBedrooms','NofBathrooms',
                  'AirConditioning','GarageSize','Pool','YearBuilt','Quality',
                   'Style','LotSize','AdjacenttoHighway'))
```

#### fit model 
```{r}
fit0 = lm(SalesPrice~.-Quality-Style, data=dataC07)
par(mfrow=c(2,2))
plot(fit0)
```

#### Find the index of influential observations which meet the following criteria: RSTUDENT > 4 or Cook’s D > 0.12 or CovRatio < 0.8

```{r}
cd<-which(cooks.distance(fit0)>0.12)
covr<-which(covratio(fit0)< 0.8)
rst<-which(rstudent(fit0)>4)
index<-sort(unique(c(cd,covr,rst)))
index
```

#### bulid table 

```{r}
DataTable <- cbind(index,rstudent(fit0)[index],
                   cooks.distance(fit0)[index],
                   covratio(fit0)[index])
colnames(DataTable) = c("ID value", "RSTUDENT", "Cook’s D","CovRatio")
DataTable
```

