---
title: 'Homework #8'
author: "Wenqiang Feng"
date: "2/25/2017"
output: html_document
---


## **1. Chapter 16 Problems 12**

#### required library
```{r message=FALSE,warning=FALSE}
library(plyr)
library(gplots)
library(asbio)
```
#### Load data
```{r}
data1612 <- read.table("./data/CH16PR12.txt")[,1:2]
colnames(data1612) = c("time_lapse", "agent")
```

## Problem 16

### problem 16.12 a
* the factor level means appear to differ
* the variability of the observations is approximately the same for all factor levels
```{r}
boxplot(time_lapse~agent, data=data1612, xlab="agent", ylab="tiem lapse")
```

### problem 16.12 b
```{r}
data1612$agent = as.factor(data1612$agent)
aov.fit = aov(time_lapse~agent, data = data1612)
fitted_value = ddply(data1612, .(agent), summarize, mean=mean(time_lapse))
colnames(fitted_value) = c("agent", "fitted value") 
fitted_value
```

### problem 16.12 c
* residuals
```{r}
residuals = aov.fit$residuals
names(residuals) = rep(1:5, each=20)
matrix(residuals, nrow=20)
```
```{r}
sum(residuals)
```

* Yes, the sum of residuals is $`r sum(residuals)`$ which is 0 at the machine error level.

### problem 16.12 d
```{r}
summary(aov.fit)
```

### problem 16.12 e

* Hypothesis:

  + $H_0$: all the $\mu_i, i= 1,\cdots,5$ are equal.
  + $H_\alpha$: not all $\mu_i, i= 1,\cdots,5$ are equal.

* Decision rules: From the above ANOVA table , we get  $F^*$ = 1107.5/7.5 = 147.6667; Now calculate F(0.90; 4, 95):
```{r}
qf(0.90, 4, 95)
```

If $F^*$ is less or equal F(0.90; 4, 95)= `r qf(0.90, 4, 95)`, conclude $H_0$, otherwise $H_\alpha$.
Based on this criteria, since $F^*$ = 147.6667 > `r qf(0.90, 4, 95)`, conclude $H_\alpha$.

### problem 16.12 f
* P value = 0

### problem 16.12 g
* Yes. There appears to be much variation in the mean time lapse among the five agents. 
* However, this variation does not necessarily indicate differences in the efficiency of operations of all five agents. It only means that at least one agent is significantly different from others.



## **1. Chapter 17 Problems 13**

### problem 17.13 a

* The intervals of some agents are not overlapped with agents of any other agents, which indicates that there is a significant difference in time lapse among the five agents. 
```{r}
plotmeans(time_lapse~agent, data=data1612, xlab="agent", ylab="time lapse", p=0.9)
```

### problem 17.13 b
* given alpha = 0.1, so the confidence level = 0.90
```{r}
tk.comparison = TukeyHSD(aov.fit, conf.level=0.90, ordered=T)
tk.comparison
```

* based on the Tukey pairwised comparison, only the mean between group 1 and 2 do not differ
```{r}
tk.comparison$agent[tk.comparison$agent[,4]>=0.05, , drop=FALSE]
```

* groups
  + group 1: agent 1 and agent 2
  + group 2: agent 3
  + group 3: agent 4
  + group 4: agent 5
  
### problem 17.13 f

* No. Based on the table above, Bonferroni has narrower intervals than the tukey procedure.


* Below is the plot of the entire result. Group pairs that intersect the dashed vertical line do not differ in time lapse mean. Here it is group pair __(1, 2)__
```{r}
plot(tk.comparison)
```


### problem 17.13 c
* MSE = 7.522632
```{r}
MSE = summary(aov.fit)[[1]][3][2,1]
MSE
```

* s{Y_bar1.} = $(MSE/20)^{0.5} = (7.522632/20)^{0.5}$ = 0.6132957
* t(0.95, 95) = 1.661052
```{r}
qt(0.95, 95)
```
```{r}
c(24.550-(MSE/20)^(0.5)*qt(0.95, 95), 24.550+(MSE/20)^(0.5)*qt(0.95, 95))
```
* 90% confidence interval = [23.53128 25.56872]
  + lower limit = Y1.mean - t(0.90, 95) * s{Y_bar1.}  = 24.55 - 1.661052 * 0.6132957 = 23.53128
  + upper limit = Y1.mean + t(0.90, 95) * s{Y_bar1.}  = 24.55 + 1.661052 * 0.6132957 = 25.56872
  
  
### problem 17.13 d
* D = mu2 - mu1 = 22.55 - 24.55 = -2
* s{Y1.} = $(MSE/(1/20 + 1/20))^{0.5} $= 0.8673
* t(0.95, 95) = `r qt(0.95, 95)`
* confidence interval: [D - t(0.90; 95) * s{Y1.}, [D - t(0.95; 95) * s{Y1.}] = [-3.441, -0.559]

* explanation: the difference is significant. Time lapse for agent 2 is significantly less than
that for agent 1.

### problem 17.13 e

* D1 = Y1 - Y3 = 12.8, 10.91794 < D1 < 14.607419
* D2 = Y1 - Y5 = -5.55, -7.43206 < D2 < -3.66794
* D3 = Y3 - Y5 = -18.35, -20.23206 < D3 < -16.46794
* s{Di} = $(MSE*2/20)^{0.5}$ = 0.8673
* B = t(1-0.1/2*3; 95) = t(0.9833; 95) = 2.158

$$ 12.8-2.158(0.8673) \leq D1 \leq 12.8+2.158(0.8673), 10.928 \leq D1 \leq 14.672 $$
$$-5.55-2.158(0.8673) \leq D2 \leq -5.55+2.158(0.8673), -7.422 \leq D2 \leq -3.678$$
$$-18.35-2.158(0.8673) \leq D3 \leq -18.35+2.158(0.8673), -20.222 \leq D3 \leq -16.478$$


* Plot shows that all differences are significant. Same conclusions are obtained as in part(a).
```{r}
D = c(12.8, -5.55, -18.35)
plot(cbind(D, 2), yaxt = "n", bty="l", ylab="", xlab="time lapse")
text(x=D, y=2.1, labels=1:3)
abline(h=2)
```

### problem 17.13 f

No, since q(0.90, 5, 95) = 3.54, while T = 2.503.


## **1. Chapter 17 Problems 18**

### problem 17.18 a

* mu1 = 24.55
* mu2 = 22.55
* mu3 = 11.75
* mu4 = 14.80
* L = (mu1+mu2)/2 - (mu3+mu4)/2 = 10.275
* s{L} = $(MSE/n)^{0.5} = (7.522632/20)^{0.5}$ = 0.6132957
* t(0.90, 95) = 1.290527
* so, L confidence interval = [L - t(0.90, 95) * s{L},  L + t(0.90, 95) * s{L}] = [9.483526 11.066474]
```{r}
mu1 = 24.55; mu2 = 22.55;mu3 = 11.75;mu4 = 14.80;n=20
L = (mu1+mu2)/2 - (mu3+mu4)/2
L
sL = (MSE/n)^0.5
sL
c(L - qt(0.90,95)*sL,L + qt(0.90, 95)*sL  )
```
explanation: the time lapse is significantly different between agents who distribute merchandise only and agents who distribute cash-value coupons only.
```{r}
mu = fitted_value[,2]
mu
MSE
qt(0.90, 95)
```


### problem 17.18 b
* D1 = mu1 - mu2 = 24.55 - 22.55 = 2.00
* D2 = mu3 - mu4 = 11.75 - 14.80 = -3.05
* L1 = (mu1+mu2)/2 - mu5 = (24.55+22.55)/2 - 30.10 = -6.55
* L2 = (mu3+mu4)/2 - mu5 = (11.75+14.80)/2 - 30.10 = -16.825
* L3 = (mu1+mu2)/2 - (mu3+mu4)/2 = (24.55+22.55)/2 - (11.75+14.80)/2 = 10.275
* 
* s{D} = 0.8673
* s{L1} = 0.7511
* s{L3} = s{D} = 0.6133
* 
* F(0.90; 4, 95) = 2.004992
* S = (4 * F(0.90; 4, 95))^0.5 = 2.831955
* Using this formula: confidence interval = [L - S * s{L},  L + t(0.90, 95) * s{L}], we obtain the * following
* confidence interval for each,

* D1: [-0.4561543, 4.4561543]
* D1: [-5.5061543, -0.5938457]
* L1: [-8.6770812, -4.4229188]
* L2: [-18.9520812, -14.6979188]
* L3: [8.5381622, 12.0118378]

explanation: difference D1 is not significant. The others are significant.
```{r}
mu = fitted_value[,2]
mu
```

### problem 17.18 c
* L = 0.25mu1 + 0.20mu2 + 0.20mu3 + 0.20mu4 + 0.15mu5 = 20.4725
* s{L} = 0.2777
* t(0.90; 95) = 1.290527

* L confidence interval: [L - t(0.90; 95) * s{L}, L + t(0.90; 95) * s{L}] = [20.114, 20.831]
```{r}
L=sum(c(0.25, 0.2, .2, .2, .15)*fitted_value[,2])
L
qt(0.90, 95)
```










