---
title: 'Homework #9'
author: "Wenqiang Feng"
date: "3/6/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### required library
```{r message=FALSE, warning=FALSE}
library(plyr)
library(lmtest)
```

### Load data
```{r}
dt <- read.table("./data/CH19PR16.txt", 
                      col.names =  c("time", "technician", "make", "disk_drive"))
dt[,2] = as.factor(dt[,2])
dt[,3] = as.factor(dt[,3])
dt[,4] = as.factor(dt[,4])
```


### problem 19.16 a
```{r}
fitted_values = ddply(dt, .(technician, make), summarize, 
      mean=mean(time))
fitted_values
```



### problem 19.16 b
```{r}
anova2way = aov(time~technician+make+technician:make, data=dt)
technician = rep(c("technician 1", "technician 2", "technician 3"), each=15)
make = rep(rep(c("make 1", "make 2", "make 3"), each=5), 3)
res=data.frame(technician, make, residuals=anova2way$residuals)
res
```



### problem 19.16 c

__Departures that can be studies__
* Nonconstancy of error variance
* Nonindependency of error items
* Outliers
* Omission of important explanatory variables
* Nonnormality of error terms


```{r}
## plot residuals vs fitted value 
plot(anova2way, which=1)
```

### problem 19.16 d
* Yes. the normality assumption appear to be reasonable here.
* __Normal probability plot__
```{r}
plot(anova2way, which=2)
```

* __Correlation coefficient__ = 0.9889246
```{r}
lm.fit = lm(time~technician+make+technician:make, data=dt)
StdErr = summary(lm.fit)$sigma
n = length(res$residuals)
ExpVals = sapply(1:n, function(k) StdErr*qnorm((k-.375)/(n+.25)) )
r = cor(ExpVals, sort(res$residuals))
r
```


### problem 19.16 e

* __Residual sequence plots__
the residual sequence plot
```{r}
plot(res$residuals, type="o", xlab="time", ylab="residuals")
par(mfrow=c(2,2))
plot(res$residuals[1:15], type="o", xlab="time order", ylab="residuals", main="Technician 1")
abline(h=0, col="gray")
plot(res$residuals[16:30], type="o", xlab="time order", ylab="residuals", main="Technician 2")
abline(h=0, col="gray")
plot(res$residuals[31:45], type="o", xlab="time order", ylab="residuals", main="Technician 3")
abline(h=0, col="gray")
```

* All three plots show a random pattern. The residuals are not serially correlated.

### problem 19.17 a
Yes. Significant factor effect exists. The lines are not parallel to each other. There is a strong interaction effect between technician and make.
```{r}
technician = dt$technician
make = dt$make
time = dt$time
interaction.plot(technician, make, time)
```

### problem 19.17 b
Yes. the interaction term accounts for most of the total variability.
```{r}
summary(anova2way)
```

### problem 19.17 c
* alternatives
  + H0: coefficients of all interaction terms  equal zero
  + Ha: not all coefficients of all interaction terms  equal zero

* F_star = 303.82/52.01 = 5.841569
* for alpha=0.01, F(0.99; 4, 36) = 3.890308 
```{r}
qf(0.99, 4, 36)
```
* decision
  + Conclude H0 if F_star > F(0.99; 4, 36)
  + Conclude Ha if F_star < F(0.99; 4, 36)
* since F_star = 5.841569 > F(0.99; 4, 36) = 3.890308, conclude Ha.
* P value = 0.000994
```{r}
summary(anova2way)[[1]]$Pr[3]
```


### problem 19.17 d
#### main effect from technician
* alternatives
  + H0: all alpha equal zero
  + Ha: not all alpha equal zero
* F_star =   12.29/52.01 = 0.2363007
* for alpha=0.01, F(0.99; 2, 36) = 5.247894
```{r}
qf(.99, 2, 36)
```
decision
  + Conclude H0 if F_star > F(0.99; 2, 36)
  + Conclude Ha if F_star < F(0.99; 2, 36)
* since F_star = 0.2363007 < F(0.99; 2, 36) = 5.247894, conclude H0.
* P value = 0.7907788
```{r}
summary(anova2way)[[1]]$Pr[1]
```

#### main effect from make
* alternatives
  + H0: all alpha equal zero
  + Ha: not all alpha equal zero
* F_star =   14.16/52.01 = 0.2722553
* for alpha=0.01, F(0.99; 2, 36) = 5.247894
```{r}
qf(.99, 2, 36)
```
* decision
    + Conclude H0 if F_star > F(0.99; 2, 36)
    + Conclude Ha if F_star < F(0.99; 2, 36)
* since F_star = 0.2722553 < F(0.99; 2, 36) = 5.247894, conclude H0.
* P value = 0.7632826
```{r}
summary(anova2way)[[1]]$Pr[2]
```

It is meaningful to test for main effect when the interaction effect is included in the model.

### problem 19.17 e
* upper bound = 1-(1-0.01)(1-0.01)(1-0.01) = 0.029701


### problem 19.17 f
* yes. the p value for interaction item is less than 0.05, which indicates that there is a significant interaction effect. While the main effects are not significant.

### problem 19.33 a 

__µ11 = [51.02908, 68.57092]__

* explanation: there is a 99% chance that the average number of minutes required by technician 1 to repair make 1 of disk drive will be located within [51.02908, 68.57092].

* MSE = 52.01111
```{r}
summary(anova2way)[[1]][3][4,]
```
* n = 5
* s{Y11.} = MSE/n = 3.225213
* t(0.995; 36) = 2.719485
```{r}
qt(.995, 36)
```
* Y11 = 59.8
```{r}
fitted_values$mean[1]
```
* µ11 = [Y11-t(0.995; 36)\*s{Y11.}, Y11+t(0.995; 36)\*s{Y11.}]
    + Y11-t(0.995; 36)\*s{Y11.} = 59.8 - 2.719485\*3.225213 = 51.02908
    + Y11+t(0.995; 36)\*s{Y11.} = 59.8 + 2.719485\*3.225213 = 68.57092
 
### problem 19.33 b

* s{D\_hat} = (2\*MSE/n)^0.5 = (2*52.01/5)^.5 = 4.56114
* D\_hat = µ22 - µ21 = 61.2 - 48.4 = 12.8
```{r}
## µ22
fitted_values$mean[5]

## µ21
fitted_values$mean[4]
```

 * t(0.995; 36) = 2.719485
```{r}
qt(0.995, 36)
```

* 99% confidence interval = [D\_hat - t(0.995; 36) * s{D\_hat}, D\_hat + t(0.995; 36) * s{D\_hat}]
    + 99% confidence interval = __[0.3960482, 25.20395]__

* explanation: there is a 99% chance that the difference between the average number of minitues needed by technician 2 and technician 1 to repaire make 1 of disc drive will be located within __[0.3960482, 25.20395]__. And technician 2 will need more time to repair the same make of disc drive.

### problem 19.33 c
* D1 = µ11 - µ12 = 12
* D2 = µ11 - µ13 = 1.4
* D3 = µ12 - µ13 = -10.6
* D4 = µ21 - µ22 = -12.8
* D5 = µ21 - µ23 = -7.8
* D6 = µ22 - µ23 = 5.0
* D7 = µ31 - µ32 = -0.6
* D8 = µ31 - µ33 = 10.6
* D9 = µ32 - µ33 = 11.2

* s{D} = 4.56114
* t(1-0.1/(2*6); 36) = 2.51104

#### Confidence intervals = [D - t(0.99-0.1/2*6; 36)*s{D}, D + t(0.99-0.1/2*6; 36)*s{D}]

```{r}
d = c(12, 1.4, -10.6, -12.6, -7.8, 5, -0.6, 10.6, 11.2)
s = 4.56114
t = 2.51104
ci = data.frame(lower = d-s*t, upper = d+s*t)
rownames(ci) = paste("D", 1:9, sep='')
ci
```

* summary:
  + for technician 1, the make 2 is the lowest
  + for technician 2, the make 1 is the lowest
  + for technician 3, the make 3 is the lowest
