---
title: "Project #1"
author: "Wenqiang Feng"
date: "2/2/2017"
output: html_document
---

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(plyr)
library(gplots)
library(asbio)
library(corrplot)
library(car)
library(MASS)
library(caret)
library(lmtest)
library(gvlma)
```
#### Load data
```{r}
C5 <- read.table("./data/APPENC05.txt")
data = C5[,2:9]
y=C5[,2]
X=C5[,3:9]
```

### Problem  1

- Stem-and-Leaf Plot
```{r}
stem(X$V3)
```
```{r}
stem(X$V4)
```
```{r}
stem(X$V5)
```
```{r}
stem(X$V6)
```
```{r}
stem(X$V7)
``` 
```{r}
stem(X$V8)
``` 
```{r}
stem(X$V9)
``` 

-  Correlation matrix
```{r}
M = cor(X)
M
```
```{r}
corrplot(M, type="upper")
```

- From the above Correlation matrix plot, we can conclude that the V3 is highly correlative with V7, V8 and V9,  V7 is highly correlative with V8. Hence i will revome V3 and V8.

### Problem 2

- linear fitting
```{r}
data2 = data[,c(1,3:6,8)]
```
```{r}
fit <- lm(V2~., data=data2)
```
```{r}
summary(fit)
```
```{r}
par(mfrow=c(2,2)) # init 4 charts in 1 panel
plot(fit)
```

- Check the data for potential influential observations
```{r}
cd<-which(cooks.distance(fit)>3.0)
covr<-which(covratio(fit)>2.0)
rst<-which(abs(rstudent(fit))>4)
index<-sort(unique(c(cd,covr,rst)))
index
```

```{r}
plot(fit, which=4)
abline(h=3.0, col="blue", lty=3)
```

- From the above analysis with the given criteria, we can conclude that obervation 32, 96, 97 are probability be the potential influential observations. 

### Problem 3 

- Check for Multicollinearity problems, and identify any variables that are involved in creating multicollinearity.
```{r}
vif_func<-function(in_frame,thresh=10,trace=T,...){

  require(fmsb)
  
  if(class(in_frame) != 'data.frame') in_frame<-data.frame(in_frame)
  
  #get initial vif value for all comparisons of variables
  vif_init<-NULL
  var_names <- names(in_frame)
  for(val in var_names){
      regressors <- var_names[-which(var_names == val)]
      form <- paste(regressors, collapse = '+')
      form_in <- formula(paste(val, '~', form))
      vif_init<-rbind(vif_init, c(val, VIF(lm(form_in, data = in_frame, ...))))
      }
  vif_max<-max(as.numeric(vif_init[,2]), na.rm = TRUE)

  if(vif_max < thresh){
    if(trace==T){ #print output of each iteration
        prmatrix(vif_init,collab=c('var','vif'),rowlab=rep('',nrow(vif_init)),quote=F)
        cat('\n')
        cat(paste('All variables have VIF < ', thresh,', max VIF ',round(vif_max,2), sep=''),'\n\n')
        }
    return(var_names)
    }
  else{

    in_dat<-in_frame

    #backwards selection of explanatory variables, stops when all VIF values are below 'thresh'
    while(vif_max >= thresh){
      
      vif_vals<-NULL
      var_names <- names(in_dat)
        
      for(val in var_names){
        regressors <- var_names[-which(var_names == val)]
        form <- paste(regressors, collapse = '+')
        form_in <- formula(paste(val, '~', form))
        vif_add<-VIF(lm(form_in, data = in_dat, ...))
        vif_vals<-rbind(vif_vals,c(val,vif_add))
        }
      max_row<-which(vif_vals[,2] == max(as.numeric(vif_vals[,2]), na.rm = TRUE))[1]

      vif_max<-as.numeric(vif_vals[max_row,2])

      if(vif_max<thresh) break
      
      if(trace==T){ #print output of each iteration
        prmatrix(vif_vals,collab=c('var','vif'),rowlab=rep('',nrow(vif_vals)),quote=F)
        cat('\n')
        cat('removed: ',vif_vals[max_row,1],vif_max,'\n\n')
        flush.console()
        }

      in_dat<-in_dat[,!names(in_dat) %in% vif_vals[max_row,1]]

      }

    return(names(in_dat))
    
    }
  
  }
```
```{r}
vif_func(in_frame=X,thresh=4,trace=T)
```

```{r}
NewData = data2[-c(index),]
fit <- lm(V2~., data=NewData)
```

```{r}
# Evaluate Collinearity
vif(fit) # variance inflation factors 
vif(fit) > 4 # problem?
```


### Problem 4 

- Check for problems with heteroscedasticity.
- Breusch-Pagan test
```{r}
lmtest::bptest(fit)
```

-  Breusch-Pagan test
```{r}
car::ncvTest(fit)  
```

With a p-value of are less than 0.05, we reject the null hypothesis (that variance of residuals is constant) and therefore infer that ther residuals are not homoscedastic. 


- Alterative method: The gvlma( ) function in the gvlma package, performs a global validation of linear model assumptions as well separate evaluations of skewness, kurtosis, and heteroscedasticity.
```{r}
gvmodel <- gvlma(fit) 
summary(gvmodel)
```
# Problem 5 

- I applied the stepwise AIC to do the variable selection.


```{r}
SWfit = stepAIC(fit, direction="both")
```

- The best model is as follows:
```{r}
summary(SWfit)
```

```{r}
par(mfrow=c(2,2)) # init 4 charts in 1 panel
plot(SWfit)
```
```{r}
gvmodel <- gvlma(SWfit) 
summary(gvmodel)
```

